Response to referees: Multi-target detection with application to cryo-electron microscopy

### 2019-05-31 InverseProblemsV2/inverse_problems_v2.tex
revision

\begin{proposition} \label{prop:uniqueness_poisson}
Under the homogeneous Poisson model, the signal $x$, the noise level $\sigma^2$ and the Poisson parameter $\gamma$ are identified uniquely \revised{from the observed autocorrelations}. 
In particular,
	\begin{align}
\gamma &= L \frac{(\gamma a_y^1) (a_y^2[1] - (a_y^1)^2)}
{\sum_{\ell=0}^{L-1} \gamma a_x^3[1,\ell] 
	+ \sum_{\ell=2}^{L-1} \gamma a_x^3[\ell,\ell+1]}, \\
\sigma^2 &= a_y^2[0] + 2\sum_{\ell = 1}^{L-1}a_y^2[\ell]-\frac{L (a^1_y)^2}{\gamma}
- (2 L - 1) (a_y^1)^2,
\end{align}
where 
\begin{align}
\gamma a_x^3[\ell_1,\ell_2]
&= a_y^3[\ell_1,\ell_2] - (a_y^1)^3
- a_y^1 \cdot \big(a_y^2[\ell_1]  
+ a_y^2[\ell_2] + a_y^2[\ell_1-\ell_2]- 3(a_y^1)^2 \big).
\end{align}

@@

This uniqueness result for two and three-dimensional signals is the basis of a popular imaging technique called coherent diffraction imaging (CDI). In CDI, an object is
illuminated with a coherent wave, and the far field diffraction pattern  is measured, corresponding to the object's Fourier magnitude~\cite{miao1999extending,shechtman2015phase}. 
If the diffraction pattern is over-sampled by at least  twice the Nyquist frequency,  the data is equivalent to the signal's second-order autocorrelation. In this context, the computational problem of recovering the signal from its second-order autocorrelation is usually referred to as \emph{phase retrieval} or the \emph{phase problem}.
Despite the uniqueness result, recently it has been shown that the problem is  ill-conditioned---at least for 2-D images~\cite{barnett2018geometry}. That is, there exist other images whose second-order autocorrelations agree up to machine precision.

@@

@@

\begin{figure}[t]
	\centering
	\includegraphics[width=1\linewidth]{heterogeneous_progressive_n12300000000_466300}
	\caption{Experiment described in Section~\ref{sec:XP1}. For a fixed noise level $\sigma = 3$ and a fixed set of $K = 3$ signals of length $L = 21$, an observation $y$ of length $N = 1.23 \cdot 10^{10}$ is generated according to~\eqref{eq:model} in the well-separated model, with fixed  occurrence probabilities. Each row corresponds to one of the signals. The last column shows evolution of the relative root mean squared error in estimating each signal, as a longer and longer subset of $y$ is observed. Red dots mark the three snapshots that are illustrated in columns 1--3: red signals are the ground truth and blue signals are the estimators.}
	\label{fig:1Dheterosignals}
\end{figure}

For the experiment depicted in Figure~\ref{fig:1Dheterosignals}, we fix $K = 3$ signals of length $L = 21$: see the three red signals in the first column. Following~\eqref{eq:model}, we generate an observation $y$ of length $12.3 \cdot 10^9$. Each of the three signals appears, respectively (and approximately), $30.0 \cdot 10^6$, $20.0 \cdot 10^6$ and $10.0 \cdot 10^6$ times in $y$ for a total of exactly $60 \cdot 10^6$ occurrences, such that at least $L-1$ zeros separate any two occurrences of any signals according to~\eqref{eq:spacing}. This is done by randomly selecting $60 \cdot 10^6$ placements in $y$, one at a time with an accept/reject rule based on the separation constraint~\eqref{eq:spacing} and locations picked so far. For each placement, one of the three signals is picked at random according to the proportions $\pi = (1/2, 1/3, 1/6)$. Then, i.i.d.\ Gaussian noise with mean zero and standard deviation $\sigma = 3$ is added, to form the observed $y$. The resulting signal-to-noise ratio of $y$
% sqrt((m_want*sum(X.^2)')/(sigma^2*n))
is about 1/9.

@@

Setting $W = L$ (as is a priori desired) is problematic because the above optimization problem appears to have numerous poor local optimizers. Thus, we first run the optimization with $W = 2L-1$. This problem appears to have few poor local optima, perhaps because the additional degrees of freedom allow for more escape directions. Since we hope the signals estimated this way correspond to the true signals zero-padded on either side to length $W$, we extract from each one a subsignal of length $L$ that has largest $\ell_2$-norm. This estimator is then used as initial iterate for~\eqref{eq:optim1D}, this time with $W = L$. We find that this procedure is reliable for a wide range of experimental parameters. To solve~\eqref{eq:optim1D}, we run the trust-region method implemented in Manopt~\cite{manopt}, which allows to treat the positivity constraints on coefficients $\hat \gamma_k$. Notice that the cost function is a polynomial in the variables, so that it is straightforward to compute it and its derivatives.

>>>>

\begin{proposition} \label{prop:uniqueness_poisson}
Under the homogeneous Poisson model, the signal $x$, the noise level $\sigma^2$ and the Poisson parameter $\gamma$ are identified uniquely \revised{from the observed autocorrelations}. 
In particular,
	\begin{align}
\gamma &= L \frac{(\gamma a_y^1) (a_y^2[1] - (a_y^1)^2)}
{\sum_{\ell=0}^{L-1} \gamma a_x^3[1,\ell] 
	+ \sum_{\ell=2}^{L-1} \gamma a_x^3[\ell,\ell+1]}, \\
\sigma^2 &= a_y^2[0] + 2\sum_{\ell = 1}^{L-1}a_y^2[\ell]-\frac{L (a^1_y)^2}{\gamma}
- (2 L - 1) (a_y^1)^2,
\end{align}
where \revised{$\gamma a_x^3$ is indeed observable since}
\begin{align}
\gamma a_x^3[\ell_1,\ell_2]
&= a_y^3[\ell_1,\ell_2] - (a_y^1)^3
- a_y^1 \cdot \big(a_y^2[\ell_1]  
+ a_y^2[\ell_2] + a_y^2[\ell_1-\ell_2]- 3(a_y^1)^2 \big).
\end{align}

@@

This uniqueness result for two and three-dimensional signals is the basis of a popular imaging technique called coherent diffraction imaging (CDI). In CDI, an object is
illuminated with a coherent wave, and the far field diffraction pattern  is measured, corresponding to the object's Fourier magnitude~\cite{miao1999extending,shechtman2015phase}. 
If the diffraction pattern is over-sampled by at least  twice the Nyquist frequency,  the data is equivalent to the signal's second-order autocorrelation. In this context, the computational problem of recovering the signal from its second-order autocorrelation is usually referred to as \emph{phase retrieval} or the \emph{phase problem}.
\revised{However, for 2-D images it has been shown recently that the problem is  ill-conditioned unless the support of the image is known exactly~\cite{barnett2018geometry}}. That is, there exist other images whose second-order autocorrelations agree up to machine precision.

@@

\revised{In this section, to assess the quality of our reconstruction, we compare against the ground truth. In a realistic setting, the ground truth is of course not available, which raises the question of how one can validate the results. A common technique used in cryo-EM is to split the data in two halves, produce two independent reconstructions based on these halves, then to compare the two reconstructions. It is clear how the same technique can be applied to our setting.}

@@

\begin{figure}[t]
	\centering
	\includegraphics[width=1\linewidth]{heterogeneous_progressive_n12300000000_143392}
	\caption{Experiment described in Section~\ref{sec:XP1}. For a fixed noise level $\sigma = 3$ and a fixed set of $K = 3$ signals of length $L = 21$, an observation $y$ of length $N = 1.23 \cdot 10^{10}$ is generated according to~\eqref{eq:model} in the well-separated model, with fixed  occurrence probabilities. Each row corresponds to one of the signals. The last column shows evolution of the relative root mean squared error in estimating each signal, as a longer and longer subset of $y$ is observed. Red dots mark the three snapshots that are illustrated in columns 1--3: red signals are the ground truth and blue signals are the estimators.}
	\label{fig:1Dheterosignals}
\end{figure}

For the experiment depicted in Figure~\ref{fig:1Dheterosignals}, we fix $K = 3$ signals of length $L = 21$: see the three red signals in the first column.
\revised{The first signal's actual support has length 11 (rather than 21), which allows us to simulate the situation in which the support of the signal is overestimated.}

Following~\eqref{eq:model}, we generate an observation $y$ of length $12.3 \cdot 10^9$. Each of the three signals appears, respectively (and approximately), $30.0 \cdot 10^6$, $20.0 \cdot 10^6$ and $10.0 \cdot 10^6$ times in $y$ for a total of exactly $60 \cdot 10^6$ occurrences, such that at least $L-1$ zeros separate any two occurrences of any signals according to~\eqref{eq:spacing}. This is done by randomly selecting $60 \cdot 10^6$ placements in $y$, one at a time with an accept/reject rule based on the separation constraint~\eqref{eq:spacing} and locations picked so far. For each placement, one of the three signals is picked at random according to the proportions $\pi = (1/2, 1/3, 1/6)$. Then, i.i.d.\ Gaussian noise with mean zero and standard deviation $\sigma = 3$ is added, to form the observed $y$. %The resulting signal-to-noise ratio of $y$
% sqrt((m_want*sum(X.^2)')/(sigma^2*n))
%is about 1/9.

@@

Setting $W = L$ (as is a priori desired) is problematic because the above optimization problem appears to have numerous poor local optimizers. Thus, we first run the optimization with $W = 2L-1$. This problem appears to have few poor local optima, perhaps because the additional degrees of freedom allow for more escape directions. Since we hope the signals estimated this way correspond to the true signals zero-padded on either side to length $W$, we extract from each one a subsignal of length $L$ that has largest $\ell_2$-norm. This estimator is then used as initial iterate for~\eqref{eq:optim1D}, this time with $W = L$. We find that this procedure is reliable for a wide range of experimental parameters. To solve~\eqref{eq:optim1D}, we run the trust-region method implemented in Manopt~\cite{manopt} \revised{from 10 different random initial guesses and keep the one with lowest cost function value.} 
\revised{Manopt} allows to treat the positivity constraints on coefficients $\hat \gamma_k$. Notice that the cost function is a polynomial in the variables, so that it is straightforward to compute it and its derivatives.

\revised{To asses reconstruction quality, we report the relative root mean squared error between the estimated signals and the ground truth (up to permutation of the $K$ signals.) For the first signal (with the overestimated support), we also translationally aligned the estimate with the ground truth because its estimation is only possible up to shift.}

### 2019-05-31 InverseProblemsV2/reponse.tex
revision

We thank the editor and the referees for reading our paper. Below, we provide a detailed, one-to-one response to all comments.

@@

\noindent \textbf{Response:} The question of statistical validation is of great importance in any scientific field, including single particle reconstruction using cryo-EM. %This question is more general that the specific method applied.
Unfortunately,  there is  no rigorous method to validate a structure that was constituted using cryo-EM. In practice, researchers in the field use several useful heuristics, such as  comparing the elucidated structure with structures that were reconstructed by other imaging modalities (e.g., X-ray crystallography), or different statistical techniques (e.g., expectation-maximization). The validation problem is even more severe in the low SNR regime (the main interest of this paper) in which visual assessment of the data might be impossible. Our method does not offer insights for this question.

While validation is beyond the scope of this paper, autocorrelation analysis provides a flexible framework to describe reach generative models of the data. We added a paragraph in the summary to explain this issue.

@@

\noindent \textbf{Response:} Section 4.4 discusses the application of our approach to 2-D images. In this section, we underscore the importance of the knowledge of the exact support for phase retrieval as was shown in [9]. In our setup, if the support cannot be estimated accurately, the solution is to consider the third-order autocorrelation. That being said, as we mention in Section 4, the support of the signal (at least in 1-D) can be determined from the second-order autocorrelation if enough data is provided. \\

@@

\noindent \textbf{Response:} We find this figure important to inspect (empirically) how many signals can be estimated simultaneously from a mix of their autocorrelations. We agree that this figure should be visualized in a colored version of the manuscript. We used a similar figure for the same purpose (but for a different problem) in a previous publication: Boumal, Nicolas, et al. ``Heterogeneous multireference alignment: A single pass approach.'' \emph{2018 52nd Annual Conference on Information Sciences and Systems (CISS)}. IEEE, 2018.

>>>>

We thank the editor and the referees for their comments on our paper. Below, we provide a detailed, one-to-one response to all comments.

@@

\noindent \textbf{Response:} We agree that statistical validation is important, in particular for single particle reconstruction using cryo-EM. %This question is more general that the specific method applied.
Several heuristics are used in cryo-EM. A common one is to split the data into two halves, reconstruct two structures independently, then compare them. 
It is easy to apply the same technique to our framework. We added a few sentences to this effect in the numerical experiments section.

%Unfortunately,  there is  no rigorous method to validate a structure that was constituted using cryo-EM. In practice, researchers in the field use several useful heuristics, such as  comparing the elucidated structure with structures that were reconstructed by other imaging modalities (e.g., X-ray crystallography), or different statistical techniques (e.g., expectation-maximization). The validation problem is even more severe in the low SNR regime (the main interest of this paper) in which visual assessment of the data might be impossible. Our method does not offer insights for this question.

%While validation is beyond the scope of this paper, autocorrelation analysis provides a flexible framework to describe reach generative models of the data. We added a paragraph in the summary to explain this issue.

%\TODO{Split the data into to halves independently and compare; checking objective and flexible model; analysis from optimization perspective of the LS }

@@

\noindent \textbf{Response:} Section 4.4 discusses the application of our approach to 2-D images. In this section, we underscore the importance of the knowledge of the exact support for phase retrieval (recovery from second-order autocorrelation) as was shown in [9]. 
We revised a couple of sentences in this section for the sake of clarity. 
If exact support knowledge is not available, our proposed solution is to consider the third-order autocorrelation.
To demonstrate it, in the revised manuscript, we changed Experiment 1. Now, half of the first signal in Figure 1 is zero, which allows us to simulate the situation in which the support of the signal is overestimated. Our algorithm still recovers the signal accurately. \\

@@

\noindent \textbf{Response:} The number of numbers we compute (as autocorrelations) is a growing function of the signal length $L$. Hence, we expect that, as $L$ grows, we should be able to estimate more and more distinct signals in the heterogeneous case. The experiment reported in Figure 2 illustrates how $K$ (the number of signals we can estimate simultaneously) grows as a function of $L$. Since to do the actual recovery we need to solve a non-convex optimization problem, this experiment is setup so that we can distinguish between the effects of statistical and computational difficulties.

We find this figure important to inspect (empirically) how many signals can be estimated simultaneously from a mix of their autocorrelations. We agree that this figure should be visualized in a colored version of the manuscript. We used a similar figure for the same purpose (but for a different problem) in a previous publication: Boumal, N., et al. ``Heterogeneous multireference alignment: A single pass approach.'' \emph{2018 52nd Annual Conference on Information Sciences and Systems (CISS)}. IEEE, 2018.
